Quantifying the Impact \\ of Performance Improvements and Cost Reductions \\ from 20 years of Light-Emitting Diode Manufacturing

### 2021-02-09 SPIE/article/main.tex
corrected error in Haitz law

\clearpage
\section{METRICS}

For instance, a highly cited and frequently updated metric is the total flux per package \cite{Liu2009,haitz2011solid,cho2017white,Fontoynont2018}. In combination with the cost per total flux, this visualization is sometimes referred to as \textit{"Haitz's Law"}, in reference to an early report on LED development by Haitz et al. \cite{haitz1999case}. However, while these metrics are often used to showcase technological progress in light-emitting diode design and manufacturing, they have limited significance XXX. Firstly, it is not desirable in all applications to increase the total flux per device beyond a certain point. Reasons for limiting the total flux per device may include lighting design considerations to reduce glare \cite{khan2015led}, device efficiency considerations to avoid electrical droop at high operating currents associated with high brightness \cite{Piprek2010} and economical considerations where multiple LED die in a single package can achieve the same brightness as a single high-brightness LED die. Secondly, even if one were to disregard these limitations, the total flux per device would only be a proxy for technological improvements in light-emitting diodes, if it was referring to single chips, instead of multi-chip paackages. 
Figure \ref{fig:haitz} shows an updated and expanded overview of best performing devices, both at the chip and package level, inspired by \textit{"Haitz's Law"}. It is evident that the historical improvement in total flux per package for single chips is not as pronounced as for multi-chip packages. While \textit{"Haitz's Law"} remains a popular visualization related to the progress in light-emitting diode design and manufacturing, care must be taken to consider its limitations.

\begin{figure} [ht]
    \begin{center}
        \includegraphics[width=\textwidth]{haitz_law_white.pdf}
    \end{center}
    \caption{Historical increase in flux for highest performing white light-emitting diode \underline{chips and (multi-chip) packages}, inspired by \textit{"Haitz's Law"}. Note that the increase in flux per single chip is not increasing as starkly as the flux for multi-chip packages. Shown are datapoints for best commercial performers from press releases, datasheets and and industry periodicals. Note the logarithmic ordinates and the black colored datapoints corresponding to the left ordinate, grey colored datapoints corresponding to the right ordinate. Sources: Data from Weinold et al. \cite{weinold2020technology}}
    \label{fig:haitz}
\end{figure}

Overall efficiency is cumulative from many sub-efficiencies caused by different physical effects \cite{schubert2018light}
In order to quantify the effect of single technological improvements on overall device performance, one must instead consider the metrics directly describing the various physical loss channels. Table \ref{tab:eff} lists the relevant sub-efficiencies considered along with mathematical definitions. The cumulative total lamp efficiency, defined as the product of all sub-efficiencies thus $\eta_L = \prod_{i=(V_f,\dots,S)} \eta_i$ XXX

The overall efficiency of a light-emitting diode is the product of the sub-efficiencies associated with an ensemble of different loss channels. Sub-efficiencies directly capture the effect of technology breakthroughs in device design and manufacturing process improvements. They thus provide the best metric to quantify technology breakthroughs.

        It describes the electrical and optical losses within the device, as well as the conversion losses in the phosphor layer. It further considers the mismatch between the device spectrum and an optimal spectrum, accounting for the wavelength dependent sensitivity of the human eye.

@@

>>>>

@@

\section{METRICS}

    Once technological breakthroughs and spillovers have been identified, their effect must be quantified. However, in solid-state lighting device performance does not only have a purely physical and economic dimension. We must consider human factors in two dimensions: First, we must consider the wavelength dependent sensitivity of the human eyes. This allows us to us to extend the physical performance dimension to an optometric performance dimension. Second, we must consider the consumer experience dimension. Related to the spectrum of the light source, this includes the perceived temperature of white light or its ability to faithfully render colors. Related to the temporal modulation of the light source, this includes unwanted flickering. Only when considering all three dimensions can the effect of a technological breakthrough be fully quantified.

    Metrics for light-emitting diodes must fulfill a number of requirements. First, the combined number of metrics must describe all dimensions of product performance introduced above. The development of a single compound metric, considering all dimensions, would be beyond the scope of this thesis. They must be widely accepted in industry. This ensures that they are reported in scientific literature and device datasheets. Metrics which have been proposed in literature but not reported by device manufacturers cannot be used to compare device performance. If neither is the case, they must at least allow for their computation from metrics which are available to us. They must further have the ability to directly describe the effect of breakthroughs on device performance. Otherwise, they must take as inputs metrics which directly describe the effect of breakthroughs on device performance. In this case, they must not require additional information which can not be fully quantified or is unavailable to us.

    For instance, a highly cited and frequently updated metric is the total luminous flux per light-emitting diode package \cite{Liu2009,haitz2011solid,cho2017white,Fontoynont2018}. In combination with the cost per total flux, this visualization is sometimes referred to as \textit{"Haitz's Law"}, in reference to an early report on LED development by Haitz et al. \cite{haitz1999case}. However, while these metrics are often used to showcase technological progress in light-emitting diode design and manufacturing, care must be taken to consider its limitations.

    Firstly, the metric today retains only limited significance as a proxy for technological development in light-emitting diodes. This is because it is not desirable to increase the total flux per device beyond a certain point in many applications. Reasons for limiting the total flux per device may include lighting design considerations to reduce glare \cite{khan2015led}, device efficiency considerations to avoid electrical droop at high operating currents associated with high brightness \cite{Piprek2010} and economical considerations where multiple LED die in a single package can achieve the same brightness as a single high-brightness LED die. Secondly, disregard these limitations, the total flux per device would only be a proxy for technological improvements in light-emitting diodes, if data was given for single light-emitting diode chips, instead of multi-chip packages. Historically, publications have sometimes failed to make this distinction, listing datapoints for both device levels in the same graph without supporting information.

    Figure \ref{fig:haitz} shows an updated and expanded overview of best performing devices, both at the chip and package level, inspired by \textit{"Haitz's Law"}. It is evident that the historical improvement in total flux per package for single chips is not as pronounced as for multi-chip packages. While \textit{"Haitz's Law"} remains a popular visualization related to the progress in light-emitting diode design and manufacturing, .

    \begin{figure} [ht]
        \begin{center}
            \includegraphics[width=\textwidth]{haitz_law_white.pdf}
        \end{center}
        \caption{Historical increase in flux for highest performing white \underline{light-emitting diode chips and (multi-chip) packages}, inspired by \textit{"Haitz's Law"}\cite{haitz1999case}. Note that flux per single chips has not increased as starkly as the flux for multi-chip packages. Shown are datapoints for best commercial performers from press releases, datasheets and and industry periodicals. Note the logarithmic ordinates and the black colored datapoints corresponding to the left ordinate, grey colored datapoints corresponding to the right ordinate. Sources: Data from Weinold et al. \cite{weinold2020technology}}
        \label{fig:haitz}
    \end{figure}

    Instead, progress in light-emitting diode technology is best described by the overall device efficiency, or lamp efficiency. This metric defined as the product of all device sub-efficiencies associated with an ensemble of different loss channels $\eta_L = \prod_{i=(V_f,\dots,S)} \eta_i$. Sub-efficiencies directly capture the effect of technology breakthroughs in device design and manufacturing process improvements. They thus provide the best metric to quantify technology breakthroughs. Table \ref{tab:eff} lists the relevant sub-efficiencies considered along with mathematical definitions.
